--- 
title: "Time Series SARIMA Model"
author: "Zivojin S. Pantic"
#date: "`r Sys.Date()`"
date: "December 2017"
site: bookdown::bookdown_site
# output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: openscapes/series
description: "Time Series SARIMA Model."
---

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
```

```{r libraries setup, echo=TRUE, message=TRUE, warning=FALSE}
# Sarima
suppressMessages(library(astsa))
suppressMessages(library(sarima))
# Arima forecast
suppressMessages(library(forecast))
# Skewness Kurtosis
suppressMessages(library(moments))
# Seasonality test
suppressMessages(library(seastests))
# Plotly
suppressMessages(library(plotly))
# Anomalies
suppressMessages(library(anomalize))
# Time series Manipulation
suppressMessages(library(xts))
# Plots
suppressMessages(library(ggplot2))
# A set of packages 
suppressMessages(library(tidyverse))
# Converting everything to everything
suppressMessages(library(tsbox))
#  'ggpubr' provides easy-to-use functions for creating and customizing 'ggplot2'- based publication ready plots
suppressMessages(library(ggpubr))
# This package has functions that plot decomposed time series data with ggplot2 
suppressMessages(library(ggplottimeseries))
# Publication ready theme
theme_set(theme_pubr())
# tables
suppressMessages(library(knitr))
suppressMessages(library(kableExtra))
# Missing data mapping
suppressMessages(library(naniar))
suppressMessages(library(pander))
```

# Abstract

The aim of this project was to construct a SARIMA time series model based on the time series of industrial production of Austria obtained from http://datamarket.com web site. After exploratory data analysis appropriate transformations, five possible models have been estimated. The best model was selected by using AIC criterion and diagnostic checking. Once SARIMA (2,1,0) (0,1,1) model was constructed and fitted it was utilized for forecasting and simulation.

# Data

```{r data}
#CSV Data -Data import from CSV file
data_csv <- read.csv('austria-qu.csv')

# Missing data mapping
vis_miss(data_csv)

#Time Series Data
data_ts<-ts(data_csv[,2],start = c(1975,4),frequency = 4)

# Conversion to tibble
data_tibble<-ts_tibbletime(data_ts)
#Extend Tibble Data
data_tibble$month <- factor(format(data_tibble$time, "%b"))
data_tibble$year <- factor(format(data_tibble$time, "%Y"))
qOrder=c('Q2','Q1','Q3','Q4')
data_tibble$quarter<-factor(data_tibble$month,labels=qOrder)

pander(data_ts)
pander(head(data_tibble))

```

# Exploratory Data Analysis

## Observations and Trend

```{r EDA}

trend <- ma(data_ts, order = 4, centre = T)
trend_ts<- as.ts(trend)
trend_tibble<-ts_tibbletime(trend_ts)

p1<-ggplot(data=data_tibble,mapping=aes(x=time,y=value))+
  geom_line(color='darkblue',size=1)+
  theme_gray()+
  labs(title = 'Time Series', x='Time', y='Observed Values')
fig1<-ggplotly(p1)
fig1
p2<-ggplot(data=trend_tibble,mapping=aes(x=time,y=value))+
  geom_line(color='darkblue', size=1)+
  theme_gray()+
  labs(title = 'Trend', x='Time', y='Trend')
fig2<-ggplotly(p2)
fig2
```

## Anomalize

```{r anomalize}
x<-as.yearqtr(data_csv$Year)
x<-as.Date(x)
y<-data_csv$Production
tbl<-tibble(x,y)
p1 <- tbl %>%
    time_decompose(y) %>%
    anomalize(remainder, alpha = 0.05, max_anoms = 0.2) %>%
    plot_anomaly_decomposition() +
    ggtitle("Anomaly Decomposition")
p2 <- tbl %>%
    time_decompose(y) %>%
    anomalize(remainder, alpha = 0.05, max_anoms = 0.2) %>%
    time_recompose() %>%
    plot_anomalies(time_recomposed = TRUE) +
    geom_line(color='darkblue')+
    ggtitle("Anomalies")
p3<-tbl %>%
  time_decompose(y) %>%
  anomalize(remainder,alpha = 0.05, max_anoms = 0.2) %>%
  time_recompose() %>%
  filter(anomaly == 'Yes')
p1
p2
p4<-tibble(x=p3$x,
           observerd=p3$observed,
           seson=p3$season,
           trend=p3$trend,
           remainder=p3$remainder,
           anomaly=p3$anomaly)
p4%>%pander()
```

## Seasonality

The WO-test gives out a TRUE if the series is seasonal and FALSE otherwise.
```{r seasonality}
summary(wo(data_ts))

# This function converts time series-class data into a data frame of decomposed time series.
df <- dts2(data_ts, type ="additive")

# plots decomposed time series into one figure
p<-ggdecompose(df)+
  theme_gray()+
  xlab("Time")+
  ylab("Index")
p
```

```{r season-plot}
p<-ggseasonplot(data_ts) +
  theme_gray()+
  ylab("Index") +
  ggtitle("Seasonal Plot")
p

p<-ggsubseriesplot(data_ts) +
  theme_gray()+
  ylab("Index") +
  ggtitle("Seasonal subseries plot")
p
```

## Box Plot

```{r boxplot}
p<-ggplot(data_tibble,aes(x=factor(quarter,levels = c('Q1','Q2','Q3','Q4'),ordered=TRUE),y=value))+
  geom_boxplot()+
  theme_gray()+
  labs(title = 'BoxPlot', x='Season', y='Index')
fig<-ggplotly(p)
fig
```

```{r boxplot2, fig.show='hide'}
# To obtain stats values and  plot non-ggplot box-plot
z<-boxplot(data_ts~cycle(data_ts),xlab='Quarter',col='beige', 
           ylab='Industrial Production Index', ylim=c(50,140), axes=FALSE)
headsv<-c('Q1','Q2','Q3','Q4')
headsh<-c('Minimum','25th Percentile','Median','75th Percentile','Maximum')
zdf<-data.frame(z$stats)
rownames(zdf)<-headsh
colnames(zdf)<-headsv
zdf%>%
  kable%>%
  kable_styling(bootstrap_options = c("responsive"), full_width = F, position = "left")%>%
  column_spec(1:1, bold = T) 
```

# Data Transformation

## Differencing

```{r differencing}
diffdata<-diff(diff(data_ts,4))
tibble_diff<-ts_tibbletime(diffdata)
p<-ggplot(data=tibble_diff,mapping=aes(x=time,y=value))+ 
  geom_line(color='darkblue')+
  theme_gray()+
  labs(title = 'Differenced Series', x='Time', y='Diffrenced Series')
fig<-ggplotly(p)
fig

```

## Lambda Parameter
```{r lambda}
lmbd<-BoxCox.lambda(data_ts, method=c("loglik"), lower=-1,upper=1)
pander(cat('lambda=',lmbd))
```

## Box-Cox Transformation

```{r boxcox}
trdata<-forecast::BoxCox(data_ts,lmbd)
difftrdata<-diff(diff(trdata,4))
tibble_difftr<-ts_tibbletime(difftrdata)
p<-ggplot(data=tibble_difftr,mapping=aes(x=time,y=value))+ 
  geom_line(color='darkblue')+
  theme_gray()+
  labs(title = 'Differenced and Transformed Series', x='Time', y='Differenced and Transformed Series')
fig<-ggplotly(p)
fig
```

## ACF and PACF

```{r acf-pacf}
p <- ggAcf(difftrdata)+
  theme_gray()+
  labs(title = 'ACF')
fig<-ggplotly(p)
fig

p <- ggPacf(difftrdata)+
  theme_gray()+
  labs(title = 'PACF')
fig<-ggplotly(p)
fig
```

# Fitting

Based on Summary of rules for identifying ARIMA models https://people.duke.edu/~rnau/arimrule.htm

```{r fitting}

fit1<-Arima(data_ts,order=c(0,1,0), seasonal=c(0,1,1),lambda=lmbd)
fit2<-Arima(data_ts,order=c(1,1,0), seasonal=c(0,1,1),lambda=lmbd)
fit3<-Arima(data_ts,order=c(0,1,1), seasonal=c(0,1,1),lambda=lmbd)
fit4<-Arima(data_ts,order=c(2,1,0), seasonal=c(0,1,1),lambda=lmbd)
fit5<-Arima(data_ts,order=c(1,1,1), seasonal=c(0,1,1),lambda=lmbd)
```

## AIC Criteria - Model Selection
```{r AIC}
aics<-c(fit1$aic,fit2$aic,fit3$aic,fit4$aic,fit5$aic)
aics
names<-c('fit1','fit2','fit3','fit4','fit5')
aicset<-data.frame(names,aics)
oset<-aicset[order(aics),]
kable(oset)%>%
  kable_styling(bootstrap_options = c("responsive"), full_width = F, position='left') %>%
  row_spec(1:1, bold = T, color = "white", background = "darkblue")
on<-which(aics == min(aics))
cat('Minimum AIC =',aics[on],' fit',on)
```

## Auto ARIMA
Auto Arima suggested model
```{r auto-Arima}
auto.arima(data_ts, d = 1, D = 1, max.p = 2, max.q = 2, max.P = 2,
  max.Q = 2, max.order = 5, seasonal = TRUE, lambda=lmbd)
```
# Confirmatiory Data Analysis

## ACF and PACF Residuals 
```{r residuals1}
resd<-fit4$residuals
p <- ggAcf(resd)+
  theme_gray()+
  labs(title = 'Residuals ACF')
fig<-ggplotly(p)
fig

p <- ggPacf(resd)+
  theme_gray()+
  labs(title = 'Residuals PACF')
fig<-ggplotly(p)
fig
```

## Residuals - Histogram

```{r residuals2}
rdf<-data.frame(resd)

p <- ggplot(rdf, aes(x=resd)) +
  geom_histogram(aes(y = ..density..), fill = "lightgrey", color="darkblue", size=.2,binwidth=.001) +
  geom_density(color = "red") +
  geom_vline(aes(xintercept=mean(resd, na.rm=T)),   
               color="red", linetype="dashed", size=.5)+
  theme_gray()+
  labs(x="")+
  labs(y="Density")+
  ggtitle("Density with Histogram Overlay")
fig <- ggplotly(p)
fig

```

## Q-Q plot, Shapiro test, skewness and kurtosis

```{r residuals3, message=FALSE, warning=FALSE}
p<-ggplot(rdf, aes(sample=resd))+
  stat_qq(color='red', shape=1)+
  stat_qq_line(color='darkblue')+
  theme_gray()+
  ggtitle("QQ Plot")
ggplotly(p)
```

```{r residuals4}
shapiro.test(resd)

sk<-skewness(resd)
kt<-kurtosis(resd)
# sk
# kt
skdf<-data.frame(Parameter=c('Skewnes','Kurtosis'),
                 Value=c(sk,kt),
                 Ideal=c(0,3),
                 Diff=c(0-sk,3-kt))
kable(skdf)%>%
  kable_styling(bootstrap_options = c("responsive"), full_width = F, position = "left")
```

## Original and Fitted Series

```{r original-fitted}
tibble_original<-ts_tibbletime(fit4$x)
tibble_fitted<-ts_tibbletime(fit4$fitted)

tibble_2<-tibble_original
tibble_2$fitted<-tibble_fitted$value
tibble_2$original<-tibble_original$value
tibble_2<-gather(tibble_2,series,val,original:fitted)

print(tibble_2)

p<-ggplot(tibble_2,aes(x=time,y=val,color=series))+
  geom_point()+
  geom_line()+
  theme(legend.position = "top")+
  theme_gray()+
  scale_color_manual(values=c("darkblue", "red"))+
  ggtitle("Original and Fitted Series")
fig <- ggplotly(p)
fig
```

# SARIMA Model Summary
```{r sarima-summary}
astsa::sarima(trdata, 2, 1, 0, P = 0, D = 1, Q = 1, S = 4, 
       details = TRUE, Model=TRUE)
```
# Forecast

```{r forecast}
fc<-forecast(fit4,12)
fcdf<-as.data.frame(fc)
kable(fcdf)%>%
  kable_styling(bootstrap_options = c("responsive"), full_width = F, position='left') 

t1<-seq(as.Date("1975/10/1"), as.Date("1991/07/1"), by = "quarter")
t2<-seq(as.Date("1991/10/1"), as.Date("1994/07/1"), by = "quarter")

df1<-data.frame(time = t1, value = data_csv$Production, series = "observations")
df2<-data.frame(time = t2, value = fcdf$`Point Forecast`, series = "forecast")
df3<-data.frame(time = t2, value = fcdf$`Lo 80`, series = "lo80")
df4<-data.frame(time = t2, value = fcdf$`Hi 80`, series = "hi80")
df5<-data.frame(time = t2, value = fcdf$`Lo 95`, series = "lo95")
df6<-data.frame(time = t2, value = fcdf$`Hi 95`, series = "hi95")

p<-ggplot() +
  geom_line(data=df1, mapping=aes(x = time, y = value,color=series), size=.4) +
  geom_line(data=df2, mapping=aes(x = time, y = value,color=series), size=.4)+
  geom_point(data=df2, mapping=aes(x = time, y = value),color='red', size=.8)+
  geom_ribbon(aes(ymin=df5$value, ymax=df6$value, x=df2$time),fill = "grey50", alpha = 0.2)+
  geom_ribbon(aes(ymin=df3$value, ymax=df4$value, x=df2$time),fill = "grey70", alpha = 0.2)+
  scale_color_manual(values=c("red", "darkblue"))+
  theme_gray()+
  labs(x="Time")+
  labs(y="Observations and Forecast")+
  ggtitle("Forecast")
fig<-ggplotly(p)
fig
```

# Time Series Cross-Validation 

Using tsCV function we measure the predictive performance of all five models

```{r cross-validation}
far1<-function(x, h){forecast(Arima(data_ts,order=c(0,1,0), seasonal=c(0,1,1),lambda=lmbd), h)}
far2<-function(x, h){forecast(Arima(data_ts,order=c(1,1,0), seasonal=c(0,1,1),lambda=lmbd), h)}
far3<-function(x, h){forecast(Arima(data_ts,order=c(0,1,1), seasonal=c(0,1,1),lambda=lmbd), h)}
far4<-function(x, h){forecast(Arima(data_ts,order=c(2,1,0), seasonal=c(0,1,1),lambda=lmbd), h)}
far5<-function(x, h){forecast(Arima(data_ts,order=c(1,1,1), seasonal=c(0,1,1),lambda=lmbd), h)}

e1<-tsCV (data_ts, far1, h = 12)
avg_e1<-mean(e1^2,na.rm=TRUE)
e2<-tsCV (data_ts, far2, h = 12)
avg_e2<-mean(e2^2,na.rm=TRUE)
e3<-tsCV (data_ts, far3, h = 12)
avg_e3<-mean(e3^2,na.rm=TRUE)
e4<-tsCV (data_ts, far4, h = 12)
avg_e4<-mean(e4^2,na.rm=TRUE)
e5<-tsCV (data_ts, far5, h = 12)
avg_e5<-mean(e5^2,na.rm=TRUE)
  
errors<-c(avg_e1,avg_e2,avg_e3,avg_e4,avg_e5) 
names<-c('MSE 1','MSE 2','MSE 3','MSE 4','MSE 5')
errorset<-data.frame(names,errors)
oerrorset<-errorset[order(errors),]

kable(oerrorset)%>%
  kable_styling(bootstrap_options = c("responsive"), full_width = F, position='left') %>%
  row_spec(1:1, bold = T, color = "white", background = "darkblue")


min.mse<-which(errors==min(errors))
cat('Minimum MSE =',errors[min.mse],' fit ',min.mse)
```

# Min MSE Model Summary
```{r min-mse-summary}
astsa::sarima(trdata, 1, 1, 0, P = 0, D = 1, Q = 1, S = 4, 
       details = TRUE, Model=TRUE)
```

# Simulation

```{r simulation}
simm<-sim_sarima(n=64, model=list(sma=fit4$coef[1],ar= fit4$coef[2],
                ma=fit4$coef[3],iorder=1, siorder=1, nseasons=4), rand.gen = rnorm)
simm_df<-data.frame(time=t1,value=simm)
p<-ggplot(simm_df,aes(x=time,y=value))+
  geom_point(color='darkblue')+
  geom_line(color='darkblue')+
  theme_gray()+
  xlab('Time') +
  ylab('Value')+
  ggtitle("Simulation")
fig<-ggplotly(p)
fig
```
